Transformers, do we really need them for detecting stress?

### 2023-06-04 paper/paper.tex
Cleanup and add abstract

\abstract{
    Here we can write an abstract of the paper.
    The abstract is a paragraph of text ranging between 70 and 150 words.
}

@@

The dataset that is used in this paper was introduced in a paper which collected posts from a social media webside called Reddit~\citep{turcan-mckeown-2019-dreaddit}.
On that website users can create posts in communities which focus on specific topics called subreddits.

The dataset consists of posts from subreddits where stressful topics are likely to be discussed in between January 1, 2017 and November 19, 2018.
These subreddits fall into one of these categories:

@@

A portion of the data was annotated using Amazon Mechanical Turk, a crowdsourcing marketplace that allows individuals to outsource jobs.

Primary job for annotators was to determine if there was stress present in the sentence that they were presented.
The definition for stress was taken from the Oxford English Dictionary states that stress is `a state of mental or emotional strain or tension resulting from adverse or demanding circumstance'.
Each annotator first had to take a qualification test to ensure that they would label the sentences correctly.
In the qualification test annotators were given instructions on how to correctly label post segments.

After that, each annotator was given 5 segments which they had to annotate with one of the following labels: ``Stress'', ``Not Stress'' or ``Can't Tell''.
In addition to the qualification test, each annotator was given one of the 50 `check questions' which were labeled by creators of the dataset.
If they did not label these `check questions' correctly their annotations were not included in the final dataset.

Since posts can often be longer, which would prove to be complicated for the annotators to label, they were divided into five-sentence chunks.
Added bonus of this technique is that the data can be used to determine the specific location of stress in the post.

In total, 3,553 labeled data points were collected out of which 39\% had perfect agreement.
With 52.3\% of the data being labeled as stressful, the dataset is nearly perfectly balanced.
Labeled data was split into two subsets: train and test subset.
The train subset contains 2,838 data points and the test subset consists of 715 data points.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/embeddings}
    \caption{Relations between similar word vectors}
    \label{fig:embedding}
\end{figure}
\hfill \break
\hfill \break
\hfill \break
Logistic regression expects a fixed size input but words and sentences in natural language can be of any size.
Today, this problem is solved by using word embeddings which represent words from our dataset as multidimensional vectors.
Their advantage over some other approaches that were used in the past, such as one-hot vector representation, is that they are able to find relations between words and produce similar vectors for similar words.
The relation between similar words can be seen in figure~\ref{fig:embedding}.
In our approach we used \textbf{word2Vec}~\citep{mikolov2013efficient} embeddings which were trained on the whole training dataset and produced 300-dimensional representations.
Learned word embeddings were then used for calculating the final representation for each training example.
Here we experimented with 2 approaches:
\begin{enumerate}
    \item \label{itm:one} \textbf{Averaging of word embeddings:} summing up each representation in the input sequence and dividing it with the number of words in the sequence
    \item \textbf{TF-IDF averaging:} Term Frequency - Inverse Document Frequency is an algorithm that is used to determine an importance of the given word in the document.
    Importance is represented as a scalar value.
     This scalar value for each word is then multiplied with the word's corresponding representation.
    The final representation is then calculated in the same way as described in~\ref{itm:one}.

@@

It must be mentioned that there was some manual data cleanup needed after we got the responses from the API (labels contained an explanation behind the given result)

@@

\begin{table*}
    \centering
    \caption{Different model results}
    \label{tab:model-results}
    \begin{center}
        \begin{tabular}{|l|c|c|c|}
            \hline
            \textbf{Model} & \textbf{P} & \textbf{R} & \textbf{F1} \\ \hline
            LR + averaged word2Vec embeddings & 0.6441 & 0.7995 & 0.7134 \\ \hline
            LR + TF-IDF averaged word2Vec embeddings & 0.6859 & 0.8049 & \textbf{0.7406} \\ \hline
            DistilBERT & 0.7105 & 0.7913 & 0.7487 \\ \hline
            Chat GPT 3.5 Turbo & 0.6905 & \textbf{0.9431} & 0.7972 \\ \hline
        \end{tabular}
    \end{center}
\end{table*}

@@

In this paper we looked at different approaches for detecting stress from the Dreaddit dataset.
We continued our research on findings of the authors and tried to find out if more complex models and solutions gave much better results.
\hfill \break
\hfill \break
We believe that the simple model solves this problem well enough and that more complex and expensive transformer solutions aren't worth the extra cost.
In some rare cases, where not detecting stress could pose a problem, we can suggest trying the Chat GPT API service if the demand isn't high and the budget isn't very limited.

>>>>

\abstract{
    Sentiment analysis is a prominent branch of machine learning and, more specifically, natural language processing that is advancing from day to day.
    In the modern day and age where negative emotions such as stress are present in everyday life of most people social media allows them to share those emotions with the world at every moment.
    Using those posts from social media the dataset Dreaddit was created which collected posts from a social media site called Reddit with the purpose of detecting stress in those posts.
    In this paper we will explore different deep learning models to find which approach is most effective in detecing stress from text.
    Some models that we worked with were logistic regression along with two transformer approaches: DistilBERT and Chat GPT 3.5 Turbo.
}

@@

The dataset that is used in this paper was introduced in a paper which collected posts from a social media webside called Reddit~\citep{turcan-mckeown-2019-dreaddit}.
On that website users can create posts in communities which focus on specific topics called subreddits.
\hfill \break
\hfill \break
The dataset consists of posts from subreddits where stressful topics are likely to be discussed in between January 1, 2017 and November 19, 2018.
These subreddits fall into one of these categories:

@@

A portion of the data was annotated using Amazon Mechanical Turk, a crowdsourcing marketplace that allows individuals to outsource jobs.
\hfill \break
\hfill \break
Primary job for annotators was to determine if there was stress present in the sentence that they were presented.
The definition for stress was taken from the Oxford English Dictionary states that stress is `a state of mental or emotional strain or tension resulting from adverse or demanding circumstance'.
Each annotator first had to take a qualification test to ensure that they would label the sentences correctly.
In the qualification test annotators were given instructions on how to correctly label post segments.
After that, each annotator was given 5 segments which they had to annotate with one of the following labels: ``Stress'', ``Not Stress'' or ``Can't Tell''.
In addition to the qualification test, each annotator was given one of the 50 `check questions' which were labeled by creators of the dataset.
If they did not label these `check questions' correctly their annotations were not included in the final dataset.
Since posts can often be longer, which would prove to be complicated for the annotators to label, they were divided into five-sentence chunks.
Added bonus of this technique is that the data can be used to determine the specific location of stress in the post.
\hfill \break
\hfill \break
In total, 3,553 labeled data points were collected out of which 39\% had perfect agreement.
With 52.3\% of the data being labeled as stressful, the dataset is nearly perfectly balanced.
Labeled data was split into two subsets: train and test subset.
The train subset contains 2,838 data points and the test subset consists of 715 data points.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/embeddings}
    \caption{Relations between similar word vectors}
    \label{fig:embedding}
\end{figure}
\hfill \break
\hfill \break
\hfill \break
Logistic regression expects a fixed size input but words and sentences in natural language can be of any size.
Today, this problem is solved by using word embeddings which represent words from our dataset as multidimensional vectors.
Their advantage over some other approaches that were used in the past, such as one-hot vector representation, is that they are able to find relations between words and produce similar vectors for similar words.
The relation between similar words can be seen in figure~\ref{fig:embedding}.
In our approach we used \textbf{word2Vec}~\citep{mikolov2013efficient} embeddings which were trained on the whole training dataset and produced 300-dimensional representations.
Learned word embeddings were then used for calculating the final representation for each training example.
Here we experimented with 2 approaches:
\begin{enumerate}
    \item \label{itm:one} \textbf{Averaging of word embeddings:} summing up each representation in the input sequence and dividing it with the number of words in the sequence
    \item \textbf{TF-IDF averaging:} Term Frequency - Inverse Document Frequency is an algorithm that is used to determine an importance of the given word in the document.
    Importance is represented as a scalar value.
     This scalar value for each word is then multiplied with the word's corresponding representation.
    The final representation is then calculated in the same way as described in approach~\ref{itm:one}.

@@

It must be mentioned that there was some manual data cleanup needed after we got the responses from the API, such as explanations behind the given results that were contained in the labels.

@@

\begin{table*}
    \centering
    \caption{Different model results}
    \label{tab:model-results}
    \begin{center}
        \begin{tabular}{|l|c|c|c|}
            \hline
            \textbf{Model} & \textbf{P} & \textbf{R} & \textbf{F1} \\ \hline
            LR + averaged word2Vec embeddings & 0.6441 & 0.7995 & 0.7134 \\ \hline
            LR + TF-IDF averaged word2Vec embeddings & 0.6859 & 0.8049 & 0.7406 \\ \hline
            DistilBERT & \textbf{0.7105} & 0.7913 & 0.7487 \\ \hline
            Chat GPT 3.5 Turbo & 0.6905 & \textbf{0.9431} & \textbf{0.7972} \\ \hline
        \end{tabular}
    \end{center}
\end{table*}

@@

In this paper we looked at different approaches for detecting stress from the Dreaddit dataset.
We continued our research on findings of the authors and tried to find out if more complex models and solutions gave better results.
\hfill \break
\hfill \break
We believe that the simple model solves this problem well enough and that more complex and expensive transformer solutions aren't worth the extra cost.
In some rare cases, where not detecting stress could pose a problem, we can suggest trying the Chat GPT API service if the demand isn't high and the budget isn't limited.
